---
title: "2019-20 Global Plastic Waste"
author: "Me"
date: "22/05/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

coast_vs_waste <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/coastal-population-vs-mismanaged-plastic.csv")

mismanaged_vs_gdp <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/per-capita-mismanaged-plastic-waste-vs-gdp-per-capita.csv")

waste_vs_gdp <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/per-capita-plastic-waste-vs-gdp-per-capita.csv")

```


# Focus on Coast vs. Waste

## Questions

- Which countries have the highest proportion of coastal population?
- What is the relationship between Total and Coastal population?
- How does proportion on coastal population affect the waste amounts?
- How does proportion on coastal population affect the waste amounts considering GDP?


```{r missing_data}

dplyr::count(coast_vs_waste, Year, sort = F) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = Year,
                                         y = n)) +
  ggplot2::geom_bar(stat = "identity")


```

```{r complete_data}

coast_vs_waste %>% 
  dplyr::filter(!is.na(`Coastal population`)) %>% 
  dplyr::count(Year)

```


```{r}

coast_vs_waste %>% 
  dplyr::filter(Year == 2010) %>% 
  tidyr::drop_na(`Coastal population`, 
                 `Mismanaged plastic waste (tonnes)`,
                 `Total population (Gapminder)`) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = `Coastal population`, 
                                         y = `Total population (Gapminder)`)) +
  ggplot2::geom_point(mapping = ggplot2::aes(size = `Mismanaged plastic waste (tonnes)`,
                                             alpha = 0.3)) +
  ggplot2::scale_x_log10() +
  ggplot2::scale_y_log10()

```

The larger the coastal population, the larger the amount of mis-managed plastic waste


```{r}

coast_vs_waste %>% 
  dplyr::filter(Year == 2010) %>% 
  dplyr::mutate(ratio = `Coastal population` / `Total population (Gapminder)`) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = ratio,
                                         y = `Mismanaged plastic waste (tonnes)`)) +
  ggplot2::geom_point() + 
  ggplot2::scale_y_log10() +
  ggplot2::geom_smooth()

```

The relative amount of plastic waste drops as more of the country lives on the coast



```{r}

complete %>% 
  tidyr::drop_na() %>% 
  dplyr::mutate(ratio = `Coastal population` / `Total population (Gapminder)`) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = ratio,
                                         y = `Mismanaged plastic waste (tonnes)`)) +
  ggplot2::geom_point(mapping = ggplot2::aes(size = `Total population (Gapminder)`), 
                      show.legend = FALSE, 
                      alpha = 0.4) + 
  ggplot2::scale_y_log10() +
  ggplot2::labs(title = "Plastic waste reduces as % of population living on the coast increases") +
  ggforce::geom_mark_ellipse(ggplot2::aes(filter = `Mismanaged plastic waste (tonnes)` > 3000000, 
                                          label = Entity, 
                                          description = scales::comma(`Mismanaged plastic waste (tonnes)`)), 
                             con.colour = "red",
                             colour = "darkred") +
  ggforce::geom_mark_rect(ggplot2::aes(filter = ratio > 1.3, 
                                          label = Entity, 
                                          description = scales::comma(`Mismanaged plastic waste (tonnes)`)), 
                          label.fontsize = 9, 
                          con.colour = "green", 
                          colour = "darkgreen") +
  ggforce::geom_mark_circle(ggplot2::aes(filter = Entity == "Australia",
                                         label = Entity,
                                         description = scales::comma(`Mismanaged plastic waste (tonnes)`)),
                            con.colour = "blue",
                            colour = "darkblue") +
  ggplot2::theme_minimal()

```


# Per capita
```{r}


complete %>% 
  tidyr::drop_na() %>% 
  dplyr::mutate(ratio = `Coastal population` / `Total population (Gapminder)`,
                waste_kg = `Mismanaged plastic waste (tonnes)` * 1000,
                waste_per_capita = waste_kg / `Total population (Gapminder)`) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = ratio,
                                         y = waste_per_capita)) +
  ggplot2::geom_smooth(method = "lm") +
  ggplot2::geom_point(mapping = ggplot2::aes(size = `Total population (Gapminder)`), 
                      show.legend = FALSE, 
                      alpha = 0.4) + 
  # ggplot2::scale_y_log10() +
  ggplot2::labs(title = "Plastic waste reduces as % of population living on the coast increases", 
                subtitle = "Kilogram per person") +
  ggforce::geom_mark_ellipse(ggplot2::aes(filter = Entity == "China" | waste_per_capita > 75, 
                                          label = Entity, 
                                          description = scales::comma(waste_per_capita)), 
                             con.colour = "red",
                             colour = "darkred") +
  # ggforce::geom_mark_rect(ggplot2::aes(filter = ratio > 1.3, 
  #                                         label = Entity, 
  #                                         description = scales::comma(waste_per_capita)), 
  #                         label.fontsize = 9, 
  #                         con.colour = "green", 
  #                         colour = "darkgreen") +
  ggforce::geom_mark_circle(ggplot2::aes(filter = Entity == "Australia",
                                         label = Entity,
                                         description = scales::comma(waste_per_capita, 
                                                                     accuracy = .01)),
                            con.colour = "blue",
                            colour = "darkblue") +
  ggplot2::theme_minimal()


```



